---
title: "Takehome 3"
date: "November 2, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
format:
  html:
    toc: true
---

## Project Methodology

**Data Collection**:

We gathered data on schools in Jakarta, including their location, type, and status. We also collected information on property prices and boundaries for administrative regions in Jakarta.

**Geospatial Analysis**:

1.  Spatial Point Patterns Analysis for School Locations (Provincial and City level)
2.  LISA and HCSA Analysis of Schools (District level)
3.  Geographically weighted regression for education levels versus availability and accessibility of schools (Sub-district level)

For my team, I'll be covering a deeper dive into (3), which is GWR regression.

## Loading Packages

```{r}
pacman::p_load(sf, tidyverse, knitr, viridis, patchwork, gridExtra, kableExtra, sfdep, tmap, tmaptools, patchwork, scales, spdep, ggrepel, spatstat, GWmodel, gtsummary, olsrr, ClustGeo, corrplot, ggpubr)
```

## Dataset Preparation

There are 3 datasets we will be using:

-   `complete_data.csv`: Contains information about schools in Indonesia, including their location, type, and status.
-   `geoBoundaries-IDN-ADM2_simplified`: Contains simplified boundaries for administrative level 2 (ADM2) in Indonesia.
-   `BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA`: Contains detailed boundaries for villages (desa) in Jakarta.

## 1.0 Spatial Points Pattern Analysis of School Distribution at the Provincial and City Level

### 1.1 Data Preprocessing

```{r}
indo_schools_kaggle <- read_csv("data/aspatial/complete_data.csv")
```

```{r}
head(indo_schools_kaggle, n=20)
```

```{r}
jakarta_schools <- indo_schools_kaggle %>% 
  filter(province_name == "DKI JAKARTA",
         city_name != "Kab. Kepulauan Seribu") %>%
  filter(!is.na(long) & !is.na(lat)) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  mutate(
      school_level = case_when(
        stage %in% c("SD", "SDLB") ~ "Elementary",
        stage %in% c("SMP", "SMPLB") ~ "Junior High",
        stage %in% c("SMA", "SMK", "SMLB") ~ "Senior High",
        TRUE ~ "Special Education"
      ),
      school_type = if_else(status == "N", "Public", "Private")
    )
```

```{r}
head(jakarta_schools)
```

```{r}
adm2 <- st_read("data/geospatial", layer = "geoBoundaries-IDN-ADM2_simplified") %>%
  filter(shapeName %in% c(
      "Kota Jakarta Barat", "Kota Jakarta Pusat",
      "Kota Jakarta Selatan", "Kota Jakarta Timur",
      "Kota Jakarta Utara"
    ))
```

```{r}
batas <- st_read("data/geospatial", layer = "BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")  %>%
    select(OBJECT_ID, PROVINSI, KAB_KOTA, KECAMATAN, DESA_KELUR, 
         JUMLAH_PEN, JUMLAH_KK, LUAS_WILAY, KEPADATAN, geometry) %>%
    rename(
    province = PROVINSI,
    city = KAB_KOTA,
    district = KECAMATAN,
    subdistrict = DESA_KELUR,
    population = JUMLAH_PEN,
    total_families = JUMLAH_KK,
    area_size = LUAS_WILAY,
    popn_density = KEPADATAN
    ) %>%
    filter(!is.na(city) & city != "KEPULAUAN SERIBU") %>%
    mutate(city = case_when(
      city == "JAKARTA BARAT" ~ "Kota Jakarta Barat",
      city == "JAKARTA PUSAT" ~ "Kota Jakarta Pusat",
      city == "JAKARTA SELATAN" ~ "Kota Jakarta Selatan",
      city == "JAKARTA TIMUR" ~ "Kota Jakarta Timur",
      city == "JAKARTA UTARA" ~ "Kota Jakarta Utara",
      TRUE ~ city
    ))
```

```{r}
head(batas)
```

```{r}
district_mapping <- c(
    "TAMAN SARI" = "Kec. Taman Sari",
    "KEMAYORAN" = "Kec. Kemayoran",
    "SAWAH BESAR" = "Kec. Sawah Besar",
    "GAMBIR" = "Kec. Gambir",
    "SENEN" = "Kec. Senen",
    "CEMPAKA PUTIH" = "Kec. Cempaka Putih",
    "MENTENG" = "Kec. Menteng",
    "TANAH ABANG" = "Kec. Tanah Abang",
    "JOHAR BARU" = "Kec. Johar Baru",
    "KOJA" = "Kec. Koja",
    "PENJARINGAN" = "Kec. Penjaringan",
    "TANJUNG PRIOK" = "Kec. Tanjung Priok",
    "CILINCING" = "Kec. Cilincing",
    "GROGOL PETAMBURAN" = "Kec. Grogol Petamburan",
    "PADEMANGAN" = "Kec. Pademangan",
    "KELAPA GADING" = "Kec. Kelapa Gading",
    "CENGKARENG" = "Kec. Cengkareng",
    "TAMBORA" = "Kec. Tambora",
    "KEBON JERUK" = "Kec. Kebon Jeruk",
    "KALIDERES" = "Kec. Kali Deres",
    "PAL MERAH" = "Kec. Palmerah",
    "JATINEGARA" = "Kec. Jatinegara",
    "KEMBANGAN" = "Kec. Kembangan",
    "TEBET" = "Kec. Tebet",
    "SETIABUDI" = "Kec. Setia Budi",
    "MAMPANG PRAPATAN" = "Kec. Mampang Prapatan",
    "PASAR MINGGU" = "Kec. Pasar Minggu",
    "KEBAYORAN LAMA" = "Kec. Kebayoran Lama",
    "KEBAYORAN BARU" = "Kec. Kebayoran Baru",
    "CILANDAK" = "Kec. Cilandak",
    "PESANGGRAHAN" = "Kec. Pesanggrahan",
    "PANCORAN" = "Kec. Pancoran",
    "JAGAKARSA" = "Kec. Jagakarsa",
    "MATRAMAN" = "Kec. Matraman",
    "PULOGADUNG" = "Kec. Pulo Gadung",
    "KRAMATJATI" = "Kec. Kramat Jati",
    "PASAR REBO" = "Kec. Pasar Rebo",
    "DUREN SAWIT" = "Kec. Duren Sawit",
    "MAKASAR" = "Kec. Makasar",
    "CIRACAS" = "Kec. Ciracas",
    "CIPAYUNG" = "Kec. Cipayung",
    "CAKUNG" = "Kec. Cakung"
)

batas <- batas %>%
  mutate(district = district_mapping[district])
```

### 1.1 EDA at the Provincial Level

#### 1.1.1 Overall Distribution of Schools

```{r}
ggplot(data = jakarta_schools) +
  geom_bar(aes(x = city_name, fill = city_name)) +
  scale_fill_viridis_d(option = "plasma") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  labs(
    title = "Distribution of Schools Across Jakarta",
    x = "City",
    y = "Number of Schools"
  )
```

**Key Observations**:

-   The plot reveals significant disparity in school distribution across Jakarta
-   East Jakarta (Timur) has the highest number of schools, followed by West Jakarta (Barat) then South Jakarta (Selatan). This is consistent with the population density distribution in Jakarta.

#### 1.1.2 Distribution of School Levels

```{r}
ggplot(data = jakarta_schools) +
  geom_bar(aes(x = city_name, fill = school_level)) +
  scale_fill_viridis_d(option = "plasma") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Distribution of School Levels Across Jakarta",
    x = "City",
    y = "Number of Schools",
    fill = "School Level"
  )
```

**Key Observations**:

-   Across all municipalities, elementary schools are the most common
-   There is a relatively consistent proportion of Junior and Senior High Schools
-   Special Education schools are less common but present in all municipalities
-   While total numbers differ, the proportional distribution of school levels remains fairly consistent across cities, suggesting **planned distribution** rather than random clustering

#### 1.1.3 School Density Across Jakarta

```{r}
ggplot() +
  geom_sf(data = adm2, aes(fill = lengths(st_intersects(adm2, jakarta_schools)))) +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal() +
  labs(
    title = "School Density Across Jakarta",
    fill = "Number of\nSchools"
  )
```

##### 1.1.4 Population Density in Jakarta

```{r}
ggplot() +
  geom_sf(data = batas, aes(fill = popn_density)) +
  scale_fill_viridis_c(
    option = "plasma",
    labels = scales::comma_format()
  ) +
  theme_minimal() +
  labs(
    title = "Population Density in Jakarta",
    fill = "Population\nDensity\n(per km²)"
  )
```

**Key Observations**:

-   The density map reveals high population clusters in central Jakarta (yellow/orange areas)
-   There is an interesting mismatch with school density in some areas
-   Several high-density pockets are scattered across Jakarta

#### 1.5 Summary

-   Our preliminary EDA at the provincial level is too macro to provide detailed insights, but necessary to understand the overall distribution of schools in Jakarta

## 2.0 Additional EDA

We note that Jakarta uses

```{r}
# Convert coordinate systems for accurate distance measurement 
# UTM Zone 48S (EPSG:32748) is appropriate for Jakarta 
jakarta_schools_utm <- st_transform(jakarta_schools, 32748) 
adm2_utm <- st_transform(adm2, 32748)
batas_utm <- st_transform(batas, 32748)
```

```{r}
# For jakarta_schools_utm
glimpse(jakarta_schools_utm)


# For batas_utm
glimpse(batas_utm)


# For column names
names(jakarta_schools_utm)
names(batas_utm)

# For dimensions
dim(jakarta_schools_utm)
dim(batas_utm)
```

We'll now look at school density per subdistrict:

```{r}
# Calculate school density per subdistrict
# Join school counts with batas dataset
school_counts <- st_join(batas_utm, jakarta_schools_utm) %>%
  group_by(OBJECT_ID) %>%
  summarise(
    total_schools = n(),
    elementary_schools = sum(school_level == "Elementary"),
    junior_schools = sum(school_level == "Junior High"),
    senior_schools = sum(school_level == "Senior High"),
    public_schools = sum(school_type == "Public"),
    private_schools = sum(school_type == "Private")
  )

# Visualize school density metrics
tmap_mode("plot")
tm_shape(school_counts) +
  tm_fill("total_schools", 
          style = "quantile",
          palette = "viridis") +
  tm_borders() +
  tm_layout(title = "School Density by Subdistrict")
```

As we can see, there are some spots (in the middle and the north of Jakarta where there seems to be sparse coverage of schools. Let's cross reference our school accessibility index for more info:

```{r}
# Calculate centroids
centroids <- st_centroid(batas_utm)

# Create distance matrix
dist_matrix <- st_distance(centroids, jakarta_schools_utm)

# Find minimum distance for each centroid
min_distances <- apply(dist_matrix, 1, min)

# Add to batas
batas_utm$nearest_school_dist <- min_distances

# Visualize
tm_shape(batas_utm) +
  tm_fill("nearest_school_dist", 
          style = "quantile",
          palette = "viridis",
          title = "Distance (m)") +
  tm_borders() +
  tm_layout(title = "Distance to Nearest School")
```

### Key Observations:

1.  Spatial Distribution Patterns:

2.  Central Jakarta shows high school density (8-14 schools) but very short distances to nearest schools (4-97m)

3.  Outer regions, particularly in the east and west, have higher numbers of schools (26-65) but longer distances (294-1,014m)

4.  Some subdistricts show an interesting mismatch: high school counts but relatively long distances, suggesting uneven distribution within the subdistrict

### Accessibility implications

1.  The central areas demonstrate better accessibility (darker purple in distance map)

2.  Peripheral areas (yellow) show reduced accessibility despite sometimes having more schools

3.  This suggests that school placement might not be optimally aligned with population distribution.

## 3.0 Clustering EDA

We'll start by using hierarchical clustering. Hierarchical clustering provides a systematic, data-driven approach to identify groups of subdistricts with similar educational characteristics. This helps us select representative areas for detailed GWR analysis while ensuring we capture the full range of educational patterns in Jakarta.

```{r}
# First, create a normalized dataset with relevant variables
clustering_data <- school_counts %>%
  st_drop_geometry() %>%  # Remove geometry for numerical calculations
  mutate(
    school_density = as.numeric(total_schools),
    elementary_ratio = as.numeric(elementary_schools / total_schools),
    junior_ratio = as.numeric(junior_schools / total_schools),
    senior_ratio = as.numeric(senior_schools / total_schools),
    public_ratio = as.numeric(public_schools / total_schools)
  ) %>%
  # Select only numeric columns for clustering
  select(school_density, elementary_ratio, junior_ratio, 
         senior_ratio, public_ratio)

summary(clustering_data)
str(clustering_data)
```

We now prepare our distance matrix.

```{r}
# Normalize the data
normalized_data <- scale(clustering_data)
```

```{r}
feature_dist <- dist(normalized_data)

```

```{r}
# 4. Perform hierarchical clustering
hc <- hclust(feature_dist, method = "ward.D2")
```

Great! let's now plot the dendrogram.

```{r}
# 5. Plot dendrogram
plot(hc, main = "Hierarchical Clustering of Jakarta Subdistricts",
     xlab = "Subdistricts", ylab = "Height")
```

```{r}
# 6. Cut tree to get 5 clusters
hc_clusters <- cutree(hc, k = 5)
```

```{r}
# 7. Add clusters back to spatial data
school_counts$hc_cluster <- as.factor(hc_clusters)
```

```{r}
# 8. Visualize clusters spatially
tm_shape(school_counts) +
  tm_fill("hc_cluster", 
          palette = "Set3",
          title = "Hierarchical Clusters") +
  tm_borders() +
  tm_layout(title = "Traditional Hierarchical Clustering")

```

::: callout-note
For hclust, we use Ward's method, also called Ward's minimum variance method, is a criterion used in hierarchical clustering that focuses on minimizing the total within-cluster variance. The "D2" variant specifically refers to the squared Euclidean distance implementation. It's good for finding spatial clusters that represent similar administrative or functional areas.
:::

## 4.0 Skater Clustering

We'll now use Skater to take a glance from a MST (Minimum Spanning Tree) approach. Although we will probably use skater to define our local GWR later, and set up bandwidth parameters, we'll still be considering our previous hclust.

### 4.1.1 Create neighborhood

```{r}
clustering_data <- school_counts %>%
  st_drop_geometry() %>%
  mutate(
    school_density = as.numeric(total_schools),
    elementary_ratio = elementary_schools/total_schools,
    junior_ratio = junior_schools/total_schools,
    senior_ratio = senior_schools/total_schools,
    public_ratio = public_schools/total_schools
  ) %>%
  select(school_density, elementary_ratio, junior_ratio, 
         senior_ratio, public_ratio) %>%
  as.data.frame()

```

```{r}
jakarta_nb <- poly2nb(school_counts, queen=TRUE)
# 2. Calculate edge costs using the neighborhood structure
lcosts <- nbcosts(jakarta_nb, clustering_data)

# 3. Create spatial weights with costs
jakarta.w <- nb2listw(jakarta_nb, 
                     lcosts, 
                     style="B")

# 4. Create minimum spanning tree
jakarta.mst <- mstree(jakarta.w)
```

```{r}
# 5. Get coordinates for plotting
coords <- st_coordinates(st_centroid(school_counts))

# 6. Plot the MST
plot(st_geometry(school_counts), 
     border=gray(.5))
plot.mst(jakarta.mst, 
         coords, 
         col="blue", 
         cex.lab=0.7, 
         cex.circles=0.005, 
         add=TRUE)
```

```{r}
# 7. Compute SKATER clusters
clust5 <- skater(edges = jakarta.mst[,1:2], 
                 data = as.matrix(clustering_data), 
                 method = "euclidean", 
                 ncuts = 4)  # This will give 5 clusters

# 8. Add clusters to spatial data
school_counts$skater_cluster <- as.factor(clust5$groups)

```

```{r}
# 9. Visualize final clusters
tm_shape(school_counts) +
    tm_fill("skater_cluster", 
            palette = "Set3",
            title = "SKATER Clusters") +
    tm_borders() +
    tm_layout(title = "School Clusters in Jakarta")
```

```{r}
# 10. Check cluster sizes
print("Number of schools in each cluster:")
print(table(clust5$groups))
```

```{r}
# 11. Create cluster summary
cluster_summary <- school_counts %>%
    st_drop_geometry() %>%
    group_by(skater_cluster) %>%
    summarise(
        n_areas = n(),
        avg_schools = mean(total_schools),
        avg_elementary = mean(elementary_schools),
        avg_junior = mean(junior_schools),
        avg_senior = mean(senior_schools)
    )

print("Cluster Summary:")
print(cluster_summary)

```

### 4.1.2 Takeaways from hclust vs Skater

SKATER has a more uneven distribution with one dominant cluster (turquoise) covering most of Jakarta, compared to hierarchical clustering with a more fragmented clustering.

## 5.0 GWR for Education level

We'll first extract the education levels from our initial batas dataset.

```{r}
# First read the original file with all columns 
batas_full <- st_read("data/geospatial", layer = "BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")

batas_full_utm <- st_transform(batas_full, 32748)

# Create education metrics with original column names then rename
batas_edu_levels <- batas_full_utm %>%
  mutate(
    # Higher education ratio (diploma and above)
    higher_ed_ratio = (DIPLOMA_I + DIPLOMA_II + DIPLOMA_IV + 
                      STRATA_II + STRATA_III) / JUMLAH_PEN,
    # Basic education completion ratio
    basic_ed_ratio = (TAMAT_SD + SLTP + SLTA) / JUMLAH_PEN,
    # No education ratio 
    no_ed_ratio = (TIDAK_BELU + BELUM_TAMA) / JUMLAH_PEN
  ) %>%
  # Select and rename columns
  select(
    PROVINSI, KAB_KOTA, KECAMATAN, DESA_KELUR,
    JUMLAH_PEN, higher_ed_ratio, basic_ed_ratio, 
    no_ed_ratio, geometry
  ) %>%
  rename(
    province = PROVINSI,
    city = KAB_KOTA,
    district = KECAMATAN,
    subdistrict = DESA_KELUR,
    population = JUMLAH_PEN
  ) %>%
  # Filter and standardize city names
  filter(!is.na(city) & city != "KEPULAUAN SERIBU") %>%
  mutate(city = case_when(
    city == "JAKARTA BARAT" ~ "Kota Jakarta Barat",
    city == "JAKARTA PUSAT" ~ "Kota Jakarta Pusat",
    city == "JAKARTA SELATAN" ~ "Kota Jakarta Selatan",
    city == "JAKARTA TIMUR" ~ "Kota Jakarta Timur",
    city == "JAKARTA UTARA" ~ "Kota Jakarta Utara",
    TRUE ~ city
  ))
```

```{r}
# Now proceed with the cluster analysis
east_cluster_edu <- batas_edu_levels %>%
  filter(city == "Kota Jakarta Timur")
```

```{r}
central_cluster_edu <- batas_edu_levels %>%
  filter(city == "Kota Jakarta Pusat")
```

### 5.1.1 Cluster data prep

Now that we have our filtered education levels for our clusters, let's begin.

### 5.1.1 Education pattern visualization

```{r}
# 3. Visualize education patterns
edu_map <- tm_shape(batas_edu_levels) +
 tm_polygons(alpha = 0.1) +
 tm_shape(east_cluster_edu) +
 tm_fill(col = "higher_ed_ratio", 
         palette = "Reds",
         title = "Higher Education Ratio") +
 tm_shape(central_cluster_edu) +
 tm_fill(col = "higher_ed_ratio",
         palette = "Blues",
         title = "Higher Education Ratio") +
 tm_layout(title = "Education Levels: East vs Central Jakarta")

print(edu_map)
```

### 5.1.2 Building GWR model

```{r}
district_boundaries <- batas_utm %>%
  group_by(city, district) %>%
  summarise(
    total_population = sum(population),
    total_area = sum(area_size),
    popn_density = total_population / total_area,
    do_union = TRUE,
    .groups = 'drop'
  )

# Then do your school density analysis at district level
school_density_analysis <- district_boundaries %>%
  mutate(
    total_schools = lengths(st_intersects(., jakarta_schools_utm)),
    elementary_schools = lengths(st_intersects(., 
                               jakarta_schools_utm[jakarta_schools_utm$school_level == "Elementary",])),
    junior_schools = lengths(st_intersects(., 
                           jakarta_schools_utm[jakarta_schools_utm$school_level == "Junior High",])),
    senior_schools = lengths(st_intersects(., 
                           jakarta_schools_utm[jakarta_schools_utm$school_level == "Senior High",])),
    special_schools = lengths(st_intersects(., 
                            jakarta_schools_utm[jakarta_schools_utm$school_level == "Special Education",])),
    # Calculate school density (schools per square km)
    school_density = total_schools / total_area,
    elementary_density = elementary_schools / total_area,
    junior_high_density = junior_schools / total_area,
    senior_high_density = senior_schools / total_area,
    special_ed_density = special_schools / total_area
  )


```

```{r}
head(school_density_analysis)
```

```{r}
# 1. Basic correlation analysis
analysis_data <- data.frame(
  # Education ratios
  higher_ed_ratio = batas_edu_levels$higher_ed_ratio,
  basic_ed_ratio = batas_edu_levels$basic_ed_ratio,
  no_ed_ratio = batas_edu_levels$no_ed_ratio,
  
  # School accessibility metrics
  school_density = clustering_data$school_density,
  elementary_ratio = clustering_data$elementary_ratio,
  junior_ratio = clustering_data$junior_ratio,
  senior_ratio = clustering_data$senior_ratio,
  public_ratio = clustering_data$public_ratio,
  nearest_school_dist = batas_utm$nearest_school_dist
)
```

```{r}
# Create correlation matrix
correlation_matrix <- cor(analysis_data, use = "complete.obs")

# Visualize correlations
corrplot(correlation_matrix, 
         method = "color", 
         type = "upper",
         addCoef.col = "black", 
         number.cex = 0.7,
         tl.cex = 0.7,
         title = "Correlation Matrix: Education Levels and School Accessibility")
```

```{r}
# Create scatter plots for key relationships
p1 <- ggplot(analysis_data, aes(x = school_density, y = higher_ed_ratio)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Higher Education vs School Density",
       x = "School Density",
       y = "Higher Education Ratio") +
  theme_minimal()

p2 <- ggplot(analysis_data, aes(x = nearest_school_dist, y = higher_ed_ratio)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Higher Education vs Distance to Nearest School",
       x = "Distance to Nearest School",
       y = "Higher Education Ratio") +
  theme_minimal()

p3 <- ggplot(analysis_data, aes(x = senior_ratio, y = higher_ed_ratio)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Higher Education vs Senior School Ratio",
       x = "Senior School Ratio",
       y = "Higher Education Ratio") +
  theme_minimal()

# Arrange plots
grid.arrange(p1, p2, p3, ncol = 2)
```

```{r}
# Fit multiple regression model
model <- lm(higher_ed_ratio ~ school_density + nearest_school_dist + 
            elementary_ratio + junior_ratio + senior_ratio + public_ratio,
            data = analysis_data)

# Model diagnostics
summary(model)
ols_vif_tol(model)
```

### 5.1.3 Analysis

-   The very low R-squared (0.02692) and adjusted R-squared (0.003747) suggest that:

    -   The global model explains very little of the variation in higher education ratios

    -   GWR might be more appropriate since it can capture local variations that this global model misses

<!-- -->

-   Variable selection for GWR:

    -   Only school_density shows marginal significance (p = 0.0697)

    -   nearest_school_dist has very low significance (p = 0.6549)

    -   The ratio variables all have relatively high p-values

We will now compare simple and complex models.

## 6.0 It's time for GWR

```{r}
# First identify complete cases
complete_indices <- which(!is.na(clustering_data$senior_ratio))

# Calculate centroids first, explicitly handling the warning
centroids <- st_centroid(batas_edu_levels, of_largest_polygon = TRUE)
centroid_coords <- st_coordinates(centroids)[complete_indices,]

# Create spatial dataframes with modern CRS syntax
simple_gwr_data <- SpatialPointsDataFrame(
  coords = centroid_coords,
  data = data.frame(
    higher_ed_ratio = batas_edu_levels$higher_ed_ratio[complete_indices],
    school_density = clustering_data$school_density[complete_indices]
  ),
  proj4string = CRS("EPSG:32748")  # Modern syntax for UTM Zone 48S
)

selected_gwr_data <- SpatialPointsDataFrame(
  coords = centroid_coords,
  data = data.frame(
    higher_ed_ratio = batas_edu_levels$higher_ed_ratio[complete_indices],
    school_density = clustering_data$school_density[complete_indices],
    senior_ratio = clustering_data$senior_ratio[complete_indices]
  ),
  proj4string = CRS("EPSG:32748")  # Modern syntax for UTM Zone 48S
)

# Since the data is already in UTM Zone 48S, we don't need to transform it
simple_gwr_data_proj <- simple_gwr_data
selected_gwr_data_proj <- selected_gwr_data

```

Cool. Let's now do other variables.

```{r}

bw.fixed <- bw.gwr(
  formula = higher_ed_ratio ~ school_density, 
  data = simple_gwr_data_proj,
  approach = "CV",
  kernel = "gaussian",
  adaptive = FALSE,
  longlat = FALSE)

# Run fixed bandwidth GWR
gwr.fixed <- gwr.basic(
  formula = higher_ed_ratio ~ school_density,
  data = simple_gwr_data_proj,
  bw = bw.fixed,
  kernel = "gaussian",
  adaptive = FALSE,
  longlat = FALSE)

bw.adaptive <- bw.gwr(
  formula = higher_ed_ratio ~ school_density,
  data = simple_gwr_data_proj,
  approach = "CV",
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE)

# Run adaptive bandwidth GWR
gwr.adaptive <- gwr.basic(
  formula = higher_ed_ratio ~ school_density,
  data = simple_gwr_data_proj,
  bw = bw.adaptive,
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE)

```

```{r}
# Convert SDF into sf data.frame
gwr.sf.fixed <- st_as_sf(gwr.fixed$SDF) %>%
  st_transform(st_crs(batas_edu_levels))

gwr.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(st_crs(batas_edu_levels))

# Compare models
model_comparison <- data.frame(
  Model = c("Fixed", "Adaptive"),
  Bandwidth = c(bw.fixed, bw.adaptive),
  AICc = c(gwr.fixed$GW.diagnostic$AICc,
           gwr.adaptive$GW.diagnostic$AICc),
  R2 = c(gwr.fixed$GW.diagnostic$gw.R2,
         gwr.adaptive$GW.diagnostic$gw.R2),
  Adj_R2 = c(gwr.fixed$GW.diagnostic$gwR2.adj,
             gwr.adaptive$GW.diagnostic$gwR2.adj)
)

print("Model Diagnostics:")
print(model_comparison)
```

ok cool. Although AICC for selected is lower, it's more complex than the simple GWR. thus we should probably stick to the simple GWR so it's easier to compute and interpret.

```{r}
fixed_r2_map <- tm_shape(batas_edu_levels) +
  tm_polygons(alpha = 0.1) +
tm_shape(gwr.sf.fixed) +  
  tm_dots(col = "Local_R2",
          size = 0.1,
          border.col = "gray60",
          border.lwd = 1,
          palette = "viridis",
          title = "Local R² (Fixed)") +
  tm_layout(main.title = "Fixed Bandwidth GWR",
            main.title.size = 1)

adaptive_r2_map <- tm_shape(batas_edu_levels) +
  tm_polygons(alpha = 0.1) +
tm_shape(gwr.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          size = 0.1,
          border.col = "gray60",
          border.lwd = 1,
          palette = "viridis",
          title = "Local R² (Adaptive)") +
  tm_layout(main.title = "Adaptive Bandwidth GWR",
            main.title.size = 1)

tmap_arrange(fixed_r2_map, adaptive_r2_map, 
            asp=1, ncol=2)
```

```{r}
fixed_coef_map <- tm_shape(batas_edu_levels) +
  tm_polygons(alpha = 0.1) +
tm_shape(gwr.sf.fixed) +  
  tm_dots(col = "school_density",
          size = 0.1,
          border.col = "gray60",
          border.lwd = 1,
          palette = "RdBu",
          title = "Coefficient (Fixed)") +
  tm_layout(main.title = "School Density Effect (Fixed)",
            main.title.size = 1)

adaptive_coef_map <- tm_shape(batas_edu_levels) +
  tm_polygons(alpha = 0.1) +
tm_shape(gwr.sf.adaptive) +  
  tm_dots(col = "school_density",
          size = 0.1,
          border.col = "gray60",
          border.lwd = 1,
          palette = "RdBu",
          title = "Coefficient (Adaptive)") +
  tm_layout(main.title = "School Density Effect (Adaptive)",
            main.title.size = 1)

tmap_arrange(fixed_coef_map, adaptive_coef_map, 
            asp=1, ncol=2)
```

```{r}
# Summary of Local R2
cat("\nLocal R² Summary (Fixed):\n")
print(summary(gwr.fixed$SDF$Local_R2))
cat("\nLocal R² Summary (Adaptive):\n")
print(summary(gwr.adaptive$SDF$Local_R2))

# Summary of Coefficients
cat("\nCoefficient Summary (Fixed):\n")
print(summary(gwr.fixed$SDF$school_density))
cat("\nCoefficient Summary (Adaptive):\n")
print(summary(gwr.adaptive$SDF$school_density))
```

We'll now export the data.

```{r}
# Transform coordinates back to WGS84 for mapping
simple_gwr_data_wgs84 <- spTransform(simple_gwr_data_proj, CRS("EPSG:4326"))

gwr_results <- list(
  # Model results
  fixed = list(
    model = gwr.fixed,
    bandwidth = bw.fixed,
    local_R2 = gwr.fixed$SDF$Local_R2,
    coefficients = gwr.fixed$SDF$school_density,
    diagnostics = gwr.fixed$GW.diagnostic
  ),
  
  adaptive = list(
    model = gwr.adaptive,
    bandwidth = bw.adaptive,
    local_R2 = gwr.adaptive$SDF$Local_R2,
    coefficients = gwr.adaptive$SDF$school_density,
    diagnostics = gwr.adaptive$GW.diagnostic
  ),
  
  # Spatial data with both coordinate systems
  spatial = list(
    coordinates_utm = coordinates(simple_gwr_data_proj),  # UTM coordinates for analysis
    coordinates_wgs84 = coordinates(simple_gwr_data_wgs84),  # WGS84 coordinates for mapping
    batas = st_transform(batas_edu_levels, 4326),  # Transform boundaries to WGS84 too
    utm_crs = "EPSG:32748",
    wgs84_crs = "EPSG:4326"
  ),
  
  # Original data
  data = list(
    higher_ed_ratio = simple_gwr_data$higher_ed_ratio,
    school_density = simple_gwr_data$school_density
  )
)

# Save to RDS
saveRDS(gwr_results, "data/rds/gwr_results.rds")
```

Below is the interactive shinyapp:

<https://hxyap-is415-workbook.shinyapps.io/takehome3-gwr/>
