---
title: "Takehome 2"
date: "September 11, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
format:
  html:
    toc: true
---

# Introduction : Drug Abuse in Thailand

Drug abuse and trafficking in Thailand present a complex and evolving challenge that demands rigorous research and analysis. This study focuses on Thailand due to its unique position in the global drug trade and its recent policy shifts, offering valuable insights into the spatial and temporal dynamics of drug issues in Southeast Asia. Thailand's proximity to the Golden Triangle, the largest drug production site in Asia, coupled with its developing transportation infrastructure, has positioned the country as both a significant market and a crucial transit route for illicit substances. The COVID-19 pandemic and its aftermath have added new dimensions to this landscape, as drug syndicates rapidly adapted their operations to changing circumstances. Our research aims to examine whether key indicators of drug abuse in Thailand are independent of spatial factors, while also exploring the temporal aspects of drug trends. By analyzing data on drug prevalence, pricing, distribution methods, and law enforcement efforts over recent years, we seek to uncover patterns that may inform more effective policies and interventions. The study considers these two of critical factors:

1.  The effects of major events, such as the COVID-19 pandemic, on drug trafficking and consumption

2.  The consequences of recent legislative changes, including the 2021 Narcotics Code and the 2022 Kratom Plant Act

By examining these elements, we hope to contribute to a more nuanced understanding of drug issues in Thailand, potentially revealing insights that could be applicable to other regions facing similar challenges. This research is particularly timely as Thailand, like many countries, grapples with the post-pandemic resurgence of drug activities and seeks innovative approaches to address this persistent social and public health concern.

```{r}
pacman:: p_load(sf, tmap, tidyverse, sfdep,spdep)
```

```{r}
drug_offenses <- read_csv("data/aspatial/thai_drug_offenses_2017_2022.csv") %>%
  filter(types_of_drug_offenses == "drug_use_cases")
```

```{r}
drug_offenses <- drug_offenses %>%
  mutate(province_en = case_when(
    province_en == "buogkan" ~ "Bueng Kan",
    province_en == "Loburi" ~ "Lop Buri",
    TRUE ~ province_en
  ))
```

```{r}
adm1 <- st_read("data/geospatial/tha_adm_rtsd_itos_20210121_shp", layer = "tha_admbnda_adm1_rtsd_20220121")
```

```{r}
drug_offenses_filtered <- drug_offenses %>% select(-province_th)
adm1_filtered <- adm1 %>% select(-ADM1_TH, -ADM1ALT1TH, -ADM1ALT2TH, -ADM0_TH)
```

```{r}
jointable <- left_join(drug_offenses_filtered, adm1_filtered, by = c("province_en" = "ADM1_EN"))
```

```{r}
summary(jointable)
```

```{r}
drugoff_sf <- st_as_sf(jointable)

# Set the CRS to match adm1
st_crs(drugoff_sf) <- st_crs(adm1)

# Verify the CRS
print(st_crs(drugoff_sf))

# Check the class of the geometry column
print(class(drugoff_sf$geometry))
```

```{r}
tmap_mode("plot")

```

```{r}
plots <- list()

for (i in 2017:2022){
  p <- drugoff_sf %>%
    filter(fiscal_year == i) %>%
    tm_shape() +
    tm_polygons("no_cases")
  
  plots[[paste0("plot", i)]] <- p
}

```

```{r}
drugoff_list <- split(drugoff_sf, drugoff_sf$fiscal_year)
```

```{r, cache=TRUE}
#| cache: true
 tm_shape(drugoff_sf) +
  tm_fill("no_cases",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")
```

```{r}

```

```{r , cache=TRUE}
#| cache: true
tm_shape(drugoff_sf) +
  tm_fill("no_cases",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

```

As we can see from the equal intervals, our data is pretty skewed. Thus, we shall turn to equal quantiles to have a better visualization. With the right, we can see that actually it's only some regions that have a lot of cases, and those that do, have far more than regions than don't. revealing that our data is very skewed on a region by region basis. From this, we should look at southern and also northeast Thailand. A possible hypothesis is that the main distribution network for the Golden Triangle is located in Northeastern Thailand, with easy links to drug production and trafficking routes from Myanmar and Laos, with a secondary distribution network in Southern Thailand, allowing for potential distribution to more southerly regions, or suggesting maritime connections to the rest of Southeast Asia like Malaysia/ Indonesia.

```{r}
tmap_mode("plot")

plot_list <- list()

# Loop through the drugoff_list and create each year's plot, storing them in plot_list
for (year in names(drugoff_list)) {
  p <- tm_shape(drugoff_list[[year]]) +
      tm_fill("no_cases",
              n = 5,
              style = "quantile",
              title = "Number of Cases") +  # Set legend title
      tm_borders(alpha = 0.5) +
      tm_layout(main.title = paste("Equal Quantile Classification", year),
                main.title.size = 0.8,         # Adjust the title size
                main.title.position = c("center", "top"),  # Center the title
                frame = FALSE,                # Remove the frame around the plot
                legend.outside = TRUE,        # Place the legend outside the plot
                legend.outside.position = "right")
  
  # Store the plot in the list
  plot_list[[year]] <- p
}

# Arrange the plots in a 2x3 layout
tmap_arrange(plotlist = plot_list, ncol = 2, nrow = 3)
```

```{r}
#tm_shape(adm1_filtered) + tm_polygons() + tm_shape(ports) + tm_dots(size = 0.1)
```

# 1.0 Analyzing the impact of the Drug Offenses Act

We will be looking into 2022 as it is the year after the 2021 drug offenses act was passed.

```{r}
drugoff_list[[6]]
```

```{r}
# Generate neighbors list
wm_q_druguse22 <- drugoff_list[[6]] %>%
  mutate(nb = st_contiguity(geometry))

# Manually modify the empty neighbor set
# Suppose the empty neighbor set is at index 'empty_index' and its nearest neighbor is at index 'nearest_index'

empty_index <- 42   # Index of the geometry with the empty neighbor set
nearest_index <- 35  # Index of the nearest neighbor

# Manually add the nearest neighbor to the empty set
wm_q_druguse22$nb[[empty_index]] <- as.integer(nearest_index)

# Now, calculate the weights after modifying the neighbor list
wm_q_DUC22 <- wm_q_druguse22 %>%
  mutate(wt = st_weights(nb, style = "W"),
         .before = 1)

```

## **1.1.1 Global Measures of Spatial Autocorrelation: Moran’s I**

Moran's I is a global measure of spatial autocorrelation that helps determine whether similar values cluster together in space (positive spatial autocorrelation), disperse (negative spatial autocorrelation), or are randomly distributed. The statistic ranges from -1 to +1:

-   Values close to +1 indicate strong positive spatial autocorrelation (clustering)

-   Values close to -1 indicate strong negative spatial autocorrelation (dispersion)

-   Values near 0 suggest random spatial distribution

By using Moran's I, we can determine if the number of drug cases across Thailand's regions are spatially correlated. This will reveal if high (or low) case numbers tend to cluster together geographically. Let's perform a simple Moran's and then a monte-carlo sim to check our claim stands:

```{r}
moranI_DUC22 <- global_moran(wm_q_DUC22$no_cases, wm_q_DUC22$nb, wm_q_DUC22$wt)

glimpse(moranI_DUC22)

```

```{r}
global_moran_test(wm_q_DUC22$no_cases, wm_q_DUC22$nb, wm_q_DUC22$wt)
```

set seed

```{r}
set.seed(42069)
```

```{r}
global_moran <- global_moran_perm(wm_q_DUC22$no_cases, wm_q_DUC22$nb, wm_q_DUC22$wt, nsim = 99)

global_moran
```

Based off our Moran's test as well as our simulations, we can deduce that:

-   Since our Moran's value is positive (\>0 ) but only moderately, there is evidence for a moderate positive spatial autocorrelation, meaning that similar values tend to cluster together in space more than would be expected by chance.

    -   These results are both statistically significant with both having very small p-values that are definitely \<0.05.

    -   This is further supported by the simulation, as the observed rank of 100 out of 100 simulations means that the observed Moran's I value was higher than all simulated values under the null hypothesis of no spatial autocorrelation.

# 2.0 Local Measures

We will now confirm this with local measures.

```{r}
lisa_DUC22 <- wm_q_DUC22 %>% 
  mutate(local_moran = local_moran(no_cases, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

```{r}
tmap_mode("plot")
```

```{r}
map1 <- tm_shape(lisa_DUC22) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of no_cases",
            main.title.size = 0.8)

map2 <- tm_shape(lisa_DUC22) +
  tm_fill("p_ii",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

```{r}
lisa_sig_DUC22 <- lisa_DUC22 %>%
  filter(p_ii < 0.05)
tmap_mode("plot")
```

```{r}
tm_shape(lisa_DUC22) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig_DUC22) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

```{r}
wm_idw_DUC22 <- wm_q_druguse22 %>%
  mutate(wts = st_inverse_distance(nb, geometry,
                                   scale = 1,
                                   alpha = 1),
         .before = 1)
```

```{r}
HotColdSpotAnalysis_DUC22 <- wm_idw_DUC22 %>% 
  mutate(local_Gi = local_gstar_perm(
    no_cases, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HotColdSpotAnalysis_DUC22
```

```{r}
tmap_mode("plot")
```

```{r}
tm_shape(HotColdSpotAnalysis_DUC22) +
  tm_fill("gi_star", palette = "-RdBu", style = "cont") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

```{r}
tmap_mode("plot")
```

```{r}
tm_shape(HotColdSpotAnalysis_DUC22) + tm_fill("p_sim") + tm_borders(alpha = 0.5)
```

```{r}
tmap_mode("plot")
```

```{r}
map1 <- tm_shape(HotColdSpotAnalysis_DUC22) +
  tm_fill("gi_star", palette = "-RdBu", style = "cont") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HotColdSpotAnalysis_DUC22) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

```{r}
HCSA_sig_DUC22 <- HotColdSpotAnalysis_DUC22  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
```

```{r}
tm_shape(HotColdSpotAnalysis_DUC22) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_sig_DUC22) +
  tm_fill("gi_star", palette = "-RdBu", style = "cont") + 
  tm_borders(alpha = 0.4)
```

Interesting. So there is a cold spot here? let's find their name:\

```{r}
cold_spot <- HotColdSpotAnalysis_DUC22 %>%
  filter(p_sim < 0.05) %>%
  arrange(gi_star) %>%
  slice(1)
```

```{r}
print(cold_spot$province_en)
```

And there we have it! A cursory google search finds us that Nakhon Sawan is home the the central point of where Thailand's waterways meet to form the Chao Phraya River. It has an army unit stationed there as well as well a large wetland. A possible explanation is that the combination of few people, hard to cross wetlands, as well as not much to be done or lack of foreign activity there means there is little incentive for drug dealers to use that location for much else but drug transit. As drugs are only there in a transitory matter, all the drug cases happen elsewhere in Thailand instead.

We know from previous current knowledge that the northern hotspot can be attributed to the well known Golden Triangle. But then why is there a random hot spot at the south of Thailand?

Let's find that southern spot:

```{r}
spots_1_to_2 <- HotColdSpotAnalysis_DUC22 %>%
  filter(gi_star >= 1 & gi_star < 2) %>%
  filter(p_sim < 0.05)  # This ensures the results are statistically significant
```

```{r}
spots_1_to_2$province_en
```

Interesting. A cursory news search (<https://www.bernama.com/en/news.php?id=2304082>) suggests that this could be due to drug lords making Phattalung, a relatively landlocked state, a possible escapade location due to it's lack of accessibility (only accessible by rail and land, no airports).

## 3.0 Bonus

For our bonus, we are going to do some EDA for fun since I ran out of time due to getting hit with food poisoning :)

```{r}
drug_unf <- read_csv("data/aspatial/thai_drug_offenses_2017_2022.csv")
unique_types <- unique(drug_unf$types_of_drug_offenses)
print(unique_types)
```

```{r}
head(drug_unf)
```

```{r}
# Step 1: Filter data for 2022, remove specified case types, and reshape
wide_data <- drug_unf %>%
  filter(fiscal_year == 2022) %>%
  filter(!types_of_drug_offenses %in% c(
    "suspects_in_export_cases",
    "export_cases",
    "suspects_in_import_cases",
    "import_cases"
  )) %>%
  select(province_th, types_of_drug_offenses, no_cases) %>%
  pivot_wider(names_from = types_of_drug_offenses, values_from = no_cases, values_fill = 0)

```

```{r}
library(corrplot)
# Step 2: Create co-occurrence matrix
offense_matrix <- as.matrix(wide_data[,-1])  # Exclude the province column
# Step 2.5: Remove "_cases" from column names
colnames(offense_matrix) <- gsub("_cases$", "", colnames(offense_matrix))
# Step 3: Check for columns with zero variance
zero_var_cols <- apply(offense_matrix, 2, var) == 0
if(any(zero_var_cols)) {
  print("The following offense types have zero variance and will be removed:")
  print(names(zero_var_cols)[zero_var_cols])
  offense_matrix <- offense_matrix[, !zero_var_cols]
}

co_occurrence <- t(offense_matrix) %*% offense_matrix

# Step 5: Handle NA and Inf values in normalization
diag_values <- diag(co_occurrence)
normalization_factor <- sqrt(diag_values %*% t(diag_values))

# Replace 0 with a small number to avoid division by zero
normalization_factor[normalization_factor == 0] <- 1e-10

co_occurrence_norm <- co_occurrence / normalization_factor
```

```{r}
# Step 6: Visualize the matrix using base R
heatmap(co_occurrence_norm, 
        Rowv = NA, Colv = NA, 
        scale = "none",
        main = "Normalized Co-occurrence of Drug Offense Types",
        xlab = "Drug Offense Types", 
        ylab = "Drug Offense Types")

# Add text labels to the heatmap
text(expand.grid(1:ncol(co_occurrence_norm), 1:nrow(co_occurrence_norm)),
     labels = round(c(co_occurrence_norm), 2),
     cex = 0.7)
```

Alternatively,

```{r}
# Step 6: Prepare data for ggplot
melted_corr <- as.data.frame(co_occurrence_norm) %>%
  rownames_to_column(var = "Var1") %>%
  pivot_longer(cols = -Var1, names_to = "Var2", values_to = "value")

# Step 7: Create heatmap using ggplot2 with adjusted width and smaller text
ggplot(melted_corr, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5) + # Smaller cell text
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8), # Smaller x-axis text
    axis.text.y = element_text(size = 8), # Smaller y-axis text
    axis.title = element_blank(),
    plot.title = element_text(size = 12), # Smaller title
    legend.title = element_text(size = 10), # Smaller legend title
    legend.text = element_text(size = 8), # Smaller legend text
    aspect.ratio = 0.7  # This makes the plot wider
  ) +
  labs(title = "Normalized Co-occurrence of Drug Offense Types",
       fill = "Co-occurrence") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))

# Save the plot with a specific width and height
ggsave("wide_heatmap_small_text.png", width = 12, height = 8, units = "in", dpi = 300)
```

So essentially we can probably just combine/ rename the suspects_in columns with their respective "suspect" less column it seems :/ Cool to know.\
\
We also now know that possession in general has high correlation with drug use cases, but trafficking/production has lower correlation. I think this opens some further questions like:\
1) Where are the distribution cases located? Are they peak in the previously seen HCSA/ highlighted Moran's spots? Is the reason for our weird spots due to distributions?

2\) Where is our trafficking cases highlighted? Previously, we can recall a lull in use cases in our favourite spot Nakhon Sawan. Due to its favourable location, is it possible there is a normal amount of trafficking cases there due to it being the centre of the country of sorts?

3\) How are the cases connected? Can we map out the drug distribution networks through the general provincial data? We recall network based analysis...

4\) Hey, we did the hard work, maybe we can do Hierachial clustering or GWR (Geographically Weighted Regression?

Unfortunately it is 9pm as I write this with an upcoming group meeting, so that will be resolved next time by viewers' request\~

Thanks for reading!
